Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-15 write-ups/PLDI/paper.tex
Related work on preemption.

\paragraph{Irregular algorithms and persistent kernels}
%
There has been a lot of work on accelerating blocking irregular
algorithms using GPUs~\cite{...}, and on the \emph{persistent threads}
programming style for long-running kernels.  These approaches rely on
the occupancy-bound execution model, and aim to take ownership of the
GPU by scheduling as many workgroups as the GPU has compute units,
relying on fair scheduling between these workgroups.  These approaches
render the GPU unavailable for other tasks, and are not future-proof
because (as discussed throughout the paper) the occupancy-bound
execution model is not guaranteed by GPU programming models and is
unlikely to be guaranteed in practice by future GPU platforms.
%
\ADComment{Now say that we could update these approaches to use cooperation.}

@@

>>>>

\paragraph{Irregular algorithms and persistent kernels}

There has been a lot of work on accelerating blocking irregular
algorithms using GPUs~\cite{...}, and on the \emph{persistent threads}
programming style for long-running kernels~\cite{...}.  These
approaches rely on the occupancy-bound execution model, and aim to
take ownership of the GPU by scheduling as many workgroups as the GPU
has compute units, relying on fair scheduling between these
workgroups.  This means the GPU is unavailable for other tasks, and
the method is not future-proof because (as discussed throughout the
paper) the occupancy-bound execution model is not guaranteed by GPU
programming models and is unlikely to be guaranteed in practice by
future GPU platforms.

As we have demonstrated for a representative set of irregular
algorithms, our cooperative kernels model allows blocking algorithms
to be upgraded to run under a fair scheduler in a manner that, with
appropriate care, facilitates responsive multi-tasking.

\paragraph{Preemptive multitasking on GPUs}

Hardware support for preemptive multitasking for \nvidia GPUs has been
proposed, enabling full preemption as well as \emph{SM-draining},
whereby workgroups occupying a symmetric multiprocessor (SM; a compute
unit using our terminology) are drained until the SM becomes free for
other tasks.  A follow-up work adds the notion of SM \emph{flushing},
whereby a workgroup running on an SM can be cancelled and re-scheduled
from scratch if the workgroup has not yet committed any visible
side-effects to shared memory.  Both approaches have been evaluated
using simulators, over sets of regular GPU kernels.

The advantages of these approaches are that they do not require manual
programmer effort, and the capability to fully preempt a workgroup
guards against a rogue long-running kernel making the GPU unavailable
for other tasks.  The drawbacks are that (a) because preemption is
enabled by default, \emph{every} kernel takes a potential performance
hit, and (b) the SM draining optimisation does not help in the case of
blocking algorithms (not considered in these works), since a blocked
workgroup cannot be drained.

Our solution does not impose a performance penalty on non-cooperative
kernels, which run in an unmodified manner, and careful placement of
our new language primitives that exploit the structure of a
cooperative kernel may enable more efficient multi-tasking than
kernel-oblivious preemption.

@@

\paragraph{Cooperative multitasking}
